---
title: "Brain_volumes"
author: "Alish Palmos"
date: "05/08/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This analysis looks at MR between brain volume GWAS and Alzheimer's

#Set-up
```{r echo=FALSE}
rm(list=ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(data.table)
library(jtools)
library(knitr)
library(broom)
library(sandwich)
library(tidyverse)
library(ggplot2)
library(sgof)
library(TwoSampleMR)
library(kableExtra)
library(qvalue)

```

# Brain Volumes
```{r echo=FALSE}

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR/brain_vol/AD_Schwartzentruber/") 

#Brain vol data

brain_vol_mr <- list.files(pattern = "*.gsmr")

brain_vol_mr = lapply(brain_vol_mr, read.delim)

brain_vol_mr <- do.call(rbind.data.frame, brain_vol_mr)

brain_vol_mr <- brain_vol_mr %>%
  filter(nsnp >= 1)

brain_vol_mr_n <- count(brain_vol_mr)

```

# Calculate the number of analyses carried out
```{r echo=FALSE}

total_tests <- brain_vol_mr_n

threshold <- 0.05/total_tests

threshold <- format(threshold,scientific=FALSE)

print(threshold)
```

# Combine all to count how many outcomes we had as AD
```{r echo=FALSE}
# combine all to count how many exposures and outcomes we had
mr_df_all <- rbind(brain_vol_mr)
ad <- str_count(mr_df_all$Exposure, "AD_Schwartzentruber")
ad[is.na(ad)] <- 0
sum(ad)
```

# Combine all to count how many exposures we had as AD
```{r}
markers <- mr_df_all[- grep("AD_Schwartzentruber", mr_df_all$Exposure),]
nrow(markers)
```

# Combine full data, work out odds ratios and 95% CI
```{r}

mr_df_all$Beta_exponent <- exp(mr_df_all$bxy)
mr_df_all$LCI <- mr_df_all$Beta_exponent - (mr_df_all$se * 1.96)
mr_df_all$UCI <- mr_df_all$Beta_exponent + (mr_df_all$se * 1.96)

mr_df_all$Exposure <- gsub(".txt", "",mr_df_all$Exposure)
mr_df_all$Outcome <- gsub(".txt", "",mr_df_all$Outcome)
```

# Compute q-values and add them to the large combined data frame
```{r}
pvalues <- mr_df_all$p
qobj <- qvalue(p = pvalues)
lfdr <- qobj$lfdr
summary(qobj)
hist(qobj)
plot(qobj)

qvalues <- as.matrix(qobj$qvalues)
mr_df_all$qvalue <- cbind(qvalues)
```

# Complete cases
```{r}
mr_df_complete <- mr_df_all[complete.cases(mr_df_all), ]
```

# Adjusted p-values using SGoF method
```{r}
p <- SGoF(mr_df_complete$p)
summary(p)

plot(p)

p_list <- p$Adjusted.pvalues

mr_df_complete <- mr_df_complete[order(mr_df_complete$p),]
mr_df_complete$sgof <- p_list
```

# Adjusted p-values to be inclued in the data frame & sorted by p-value
```{r}
mr_df_complete$p.adjust <- p.adjust(p = mr_df_complete$p, method = "fdr")
```

#Get all results with AD as the outcome 
```{r} 

AD_outcome <- mr_df_complete[!grepl("AD_Schwartzentruber", mr_df_complete$Exposure),]

  AD_outcome %>%
  kbl() %>%
  kable_classic_2(full_width = F)
```

#Get all results with AD as the exposure 
```{r}
  
  AD_exposure <- mr_df_complete[!grepl("AD_Schwartzentruber", mr_df_complete$Outcome),]

  AD_exposure %>%
  kbl() %>%
  kable_classic_2(full_width = F)
```

# Sort by significant p-values and rename headers
```{r}
mr_df_filtered <- mr_df_complete %>%
  filter(qvalue < 0.05) 

mr_df_filtered <- mr_df_filtered %>%
  rename(Beta = bxy, StandardError = se, Pvalue = p, SNPs = nsnp, LowerCI = LCI, UpperCI =  UCI, AdjustedPvalue = p.adjust)

mr_df_filtered <- mr_df_filtered %>% arrange(desc(Beta_exponent))
     
mr_df_filtered$Log_Beta <- log(mr_df_filtered$Beta_exponent) 
mr_df_filtered$Log_LowerCI <- log(mr_df_filtered$LowerCI) 
mr_df_filtered$Log_UpperCI <- log(mr_df_filtered$UpperCI) 
mr_df_filtered$LCI <- mr_df_filtered$Beta - (mr_df_filtered$StandardError * 1.96)
mr_df_filtered$UCI <- mr_df_filtered$Beta + (mr_df_filtered$StandardError * 1.96)

                                        
kable(mr_df_filtered)                                

```

```{r fig.width=10, echo=FALSE}

library(scales)

mr_df_filtered <- mr_df_filtered[!grepl("AD_Schwartzentruber", mr_df_filtered$Outcome),]

f <- function(y) seq(floor(min(y)), ceiling(max(y)))

p1 <- ggplot(mr_df_filtered, aes(x=reorder(as.factor(Outcome), Beta) , y=Beta)) + 
    geom_boxplot(fill="slateblue") + coord_flip() + geom_errorbar(aes(ymax = UCI, ymin = LCI))+ ggtitle("Exposure - Alzheimer's") +
  xlab("Outcome Variable") + ylab("Beta (95% Confidence Interval)")+ geom_hline(yintercept=0, linetype="dashed", color = "red") + 
  theme_bw() + theme(text=element_text(size=16, family="Arial"))  +
    theme(panel.grid.major.x = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray41"),
        panel.grid.major.y = element_line(size = 0.1, 
                                        linetype = 'solid', 
                                        colour = "gray62")) +
  # scale_y_log10() +
  guides(color = guide_legend(reverse = TRUE)) +
  coord_flip(clip = "off", ylim = c(-55,74)) 


  
p1  

```

