---
title: "ALS_MR"
author: "Alish Palmos"
date: "09/02/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
  body{
  font-size: 12pt;
}
</style>


```{r echo=FALSE}
rm(list=ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//")
```

```{r include=FALSE}
library(data.table)
library(jtools)
library(knitr)
library(broom)
library(sandwich)
library(tidyverse)
library(ggplot2)
library(sgof)
library(TwoSampleMR)
library(qvalue)
library(kableExtra)

```

This RMarkdown shows the results from a MRWAS (Mendelian Randomization Wide Association Study). 

All currently available GWAS summary statistics of inflammatory markers were amassed over the course a few months.  

### Covid GWAS used in this study: 

Very severe respiratory confirmed covid vs. not hospitalized covid - October data release (https://www.covid19hg.org/results/) 



Sun et al: https://pubmed.ncbi.nlm.nih.gov/29875488

Suhre et al: https://pubmed.ncbi.nlm.nih.gov/28240269

Wood et al: https://pubmed.ncbi.nlm.nih.gov/23696881 

Sliz et al: https://www.ncbi.nlm.nih.gov/pubmed/?term=31217265 

Folkerson et al: http://dx.plos.org/10.1371/journal.pgen.1006706 

Ahola et al: https://pubmed.ncbi.nlm.nih.gov/27989323

Scallop et al: https://www.nature.com/articles/s42255-020-00287-2#Abs1

Bretherik et al: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7337286


Note that more studies are available for analyses. They're currently being cleaned and added as the data generation protocol progresses. 


GSMR was used as the base MR method: https://www.nature.com/articles/s41467-017-02317-2 


P-value threshold was set to 5E-6, due to lack of power in some GWAS. Thus, this study is exploratory. 


Only comparisons where 10+ SNPs were in common between the exposure and the outcome were included. This is routine MR practice. 


Bonferroni p-value threshold is stated below. 

#PROTEIN -> AD
#SUN

```{r echo=FALSE}

#SUN

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//sun_mr/output2021/")

sun_list <- list.files(pattern = "*.gsmr")

sun_mr = lapply(sun_list, read.delim)

sun_mr <- do.call(rbind.data.frame, sun_mr)

sun_mr$study <- "Sun"

sun_mr$file_name <- rep(sun_list, each=2)

sun_mr <- sun_mr %>%
  filter(nsnp >= 1)

# sun_mr$Exposure <- sub("^(.*)[.].*", "\\1",sun_mr$Exposure)
# sun_mr$Exposure <- sub("^(.*)[.].*", "\\1",sun_mr$Exposure)
# sun_mr$Exposure <- sub("^(.*)[.].*", "\\1",sun_mr$Exposure)
# sun_mr$Exposure <- sub("^(.*)[.].*", "\\1",sun_mr$Exposure)
# 
# sun_mr$Outcome <- sub("^(.*)[.].*", "\\1",sun_mr$Outcome)
# sun_mr$Outcome <- sub("^(.*)[.].*", "\\1",sun_mr$Outcome)
# sun_mr$Outcome <- sub("^(.*)[.].*", "\\1",sun_mr$Outcome)
# sun_mr$Outcome <- sub("^(.*)[.].*", "\\1",sun_mr$Outcome)

sun_n <- count(sun_mr)



```

#SUHRE

```{r echo=FALSE}

#SUHRE
setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//suhr_mr/outputs2021/")

suhre_list <- list.files(pattern = "*.gsmr")

suhre_mr = lapply(suhre_list, read.delim)

suhre_mr <- do.call(rbind.data.frame, suhre_mr)

suhre_mr$study <- "Suhre"

suhre_mr$file_name <- rep(suhre_list, each=2)

suhre_mr <- suhre_mr %>%
  filter(nsnp >= 1)

# suhre_mr$Exposure <- gsub("\\..*","",suhre_mr$Exposure)
# suhre_mr$Exposure <- gsub("(.*)_.*", "\\1",suhre_mr$Exposure)
# suhre_mr$Exposure <- gsub("(.*)_.*", "\\1",suhre_mr$Exposure)
# suhre_mr$Exposure <- gsub("(.*)_.*", "\\1",suhre_mr$Exposure)
# suhre_mr$Exposure <- gsub("Suhre_Renamed/", "\\1",suhre_mr$Exposure)


suhre_n <- count(suhre_mr)
```

#FOLK 

```{r echo=FALSE}

#FOLK 

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//folk_mr/output2021/")

folk_list <- list.files(pattern = "*.gsmr")

folk_mr = lapply(folk_list, read.delim)

folk_mr <- do.call(rbind.data.frame, folk_mr)

folk_mr$study <- "Folk"

folk_mr$file_name <- rep(folk_list, each=2)

folk_mr <- folk_mr %>%
  filter(nsnp >= 1)

folk_n <- count(folk_mr)

```

#SLIZ 

```{r echo=FALSE}

#SLIZ 

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//sliz_mr/output2021/")

sliz_list <- list.files(pattern = "*.gsmr")

sliz_mr = lapply(sliz_list, read.delim)

sliz_mr <- do.call(rbind.data.frame, sliz_mr)

sliz_mr$study <- "Sliz"

sliz_mr$file_name <- rep(sliz_list, each=2)


sliz_mr <- sliz_mr %>%
  filter(nsnp >= 1)

sliz_n <- count(sliz_mr)

```

#WOOD

```{r echo=FALSE}

#WOOD

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//wood_mr/output2021/")

wood_list <- list.files(pattern = "*.gsmr")

wood_mr = lapply(wood_list, read.delim)

wood_mr <- do.call(rbind.data.frame, wood_mr)

wood_mr$study <- "Wood"

wood_mr$file_name <- rep(wood_list, each=2)

wood_mr <- wood_mr %>%
  filter(nsnp >= 1)

wood_n <- count(wood_mr)

```

#AHOL 

```{r echo=FALSE}

#AHOL 

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//infl_mr/output2021/")

ahol_list <- list.files(pattern = "*.gsmr")

ahol_mr = lapply(ahol_list, read.delim)

ahol_mr <- do.call(rbind.data.frame, ahol_mr)

ahol_mr$study <- "Ahol"

ahol_mr$file_name <- rep(ahol_list, each=2)

ahol_mr <- ahol_mr %>%
  filter(nsnp >= 1)

ahol_n <- count(ahol_mr)

```

#SCALLOP

```{r echo=FALSE}

#SCALLOP

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//scal_mr/output2021/")

scal_list <- list.files(pattern = "*.gsmr")

scal_mr = lapply(scal_list, read.delim)

scal_mr <- do.call(rbind.data.frame, scal_mr)

scal_mr$study <- "Scal"

scal_mr$file_name <- rep(scal_list, each=2)

scal_mr <- scal_mr %>%
  filter(nsnp >= 1)

scal_n <- count(scal_mr)

```

#HILL

```{r echo=FALSE}

#HILL

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//hill_mr/output2021/")

hill_list <- list.files(pattern = "*.gsmr")

hill_mr = lapply(hill_list, read.delim)

hill_mr <- do.call(rbind.data.frame, hill_mr)

hill_mr$study <- "Hill"

hill_mr$file_name <- rep(hill_list, each=2)


hill_mr <- hill_mr %>%
  filter(nsnp >= 1)

hill_n <- count(hill_mr)

```

#HOGL

```{r echo=FALSE}

#HOGL

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//hogl_mr/output2021/")

hogl_list <- list.files(pattern = "*.gsmr")

hogl_mr = lapply(hogl_list, read.delim)

hogl_mr <- do.call(rbind.data.frame, hogl_mr)

hogl_mr$study <- "Hogl"

hogl_mr$file_name <- rep(hogl_list, each=2)


hogl_mr <- hogl_mr %>%
  filter(nsnp >= 1)

hogl_n <- count(hogl_mr)

```

#ENROTH

```{r echo=FALSE}

#ENROTH

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//enroth_mr//output2021/")

enroth_list <- list.files(pattern = "*.gsmr")

enroth_mr = lapply(enroth_list, read.delim)

enroth_mr <- do.call(rbind.data.frame, enroth_mr)

enroth_mr$study <- "Enroth"

enroth_mr$file_name <- rep(enroth_list, each=2)


enroth_mr <- enroth_mr %>%
  filter(nsnp >= 1)

enroth_n <- count(enroth_mr)

```

#Count the number of analyses done
```{r echo=FALSE}
total_tests <- sum(ahol_n$n, folk_n$n, scal_n$n, sliz_n$n, suhre_n$n, sun_n$n, wood_n$n, hill_n$n, hogl_n$n, enroth_n$n)

threshold <- 0.05/total_tests

threshold <- format(threshold,scientific=FALSE)

print(threshold)
```


# Combine all to count how many exposures we had as proteins
```{r echo=FALSE}
# combine all to count how many exposures and outcomes we had
mr_df_all <- rbind(folk_mr, sliz_mr, suhre_mr, sun_mr, wood_mr, ahol_mr, scal_mr, hill_mr, hogl_mr, enroth_mr)
als <- str_count(mr_df_all$Exposure, "AD_Schwartzentruber")
als[is.na(als)] <- 0
sum(als)
```

# Combine all to count how many exposures we had as ALS
```{r}
markers <- mr_df_all[- grep("AD_Schwartzentruber", mr_df_all$Exposure),]
nrow(markers)
```

# Combine full data, work out odds ratios and 95% CI
```{r}

mr_df_all$Beta_exponent <- exp(mr_df_all$bxy)
mr_df_all$LCI <- mr_df_all$Beta_exponent - (mr_df_all$se * 1.96)
mr_df_all$UCI <- mr_df_all$Beta_exponent + (mr_df_all$se * 1.96)

mr_df_all$Exposure <- gsub(".txt", "",mr_df_all$Exposure)
mr_df_all$Outcome <- gsub(".txt", "",mr_df_all$Outcome)
```

# Compute q-values and add them to the large combined data frame
```{r}
pvalues <- mr_df_all$p
qobj <- qvalue(p = pvalues)
lfdr <- qobj$lfdr
summary(qobj)
hist(qobj)
plot(qobj)

qvalues <- as.matrix(qobj$qvalues)
mr_df_all$qvalue <- cbind(qvalues)
```

# Adjusted p-values using SGoF method
```{r}
p <- SGoF(mr_df_all$p)
summary(p)

plot(p)

p_list <- p$Adjusted.pvalues

mr_df_all <- mr_df_all[order(mr_df_all$p),]
mr_df_all$sgof <- p_list
```

# All effects AD exposure

```{r}

AD_outcome <- mr_df_all[!grepl("AD_Schwartzentruber", mr_df_all$Exposure),]

  AD_outcome %>%
  kbl() %>%
  kable_classic_2(full_width = F)

```

# Sort by significant p-values and filter
```{r}
mr_df_filtered <- AD_outcome %>%
  filter(qvalue < 0.05) 

mr_df_filtered <- mr_df_filtered %>%
  rename(Beta = bxy, StandardError = se, Pvalue = p, SNPs = nsnp, LowerCI = LCI, UpperCI =  UCI)

mr_df_filtered <- mr_df_filtered %>% arrange(desc(Beta_exponent))
     
mr_df_filtered$Log_Beta <- log(mr_df_filtered$Beta_exponent) 
mr_df_filtered$Log_LowerCI <- log(mr_df_filtered$LowerCI) 
mr_df_filtered$Log_UpperCI <- log(mr_df_filtered$UpperCI) 
                                        
kable(mr_df_filtered)                                

```

# Create a figure with AD as the outcome 
```{r fig.width=20, fig.height=40, echo=FALSE}

mr_df_chart <- mr_df_filtered %>%
  filter(qvalue < 0.0001) 

mr_df_chart <- mr_df_chart[complete.cases(mr_df_chart), ]

#Tidy up the names of some proteins 
mr_df_chart$Exposure <- sub("^(.*)[.].*", "\\1",mr_df_chart$Exposure)
mr_df_chart$Exposure <- sub("^(.*)[.].*", "\\1",mr_df_chart$Exposure)
mr_df_chart$Exposure <- sub("^(.*)[.].*", "\\1",mr_df_chart$Exposure)
mr_df_chart$Exposure <- sub("^(.*)[.].*", "\\1",mr_df_chart$Exposure)

mr_df_chart$Outcome <- sub("^(.*)[.].*", "\\1",mr_df_chart$Outcome)
mr_df_chart$Outcome <- sub("^(.*)[.].*", "\\1",mr_df_chart$Outcome)
mr_df_chart$Outcome <- sub("^(.*)[.].*", "\\1",mr_df_chart$Outcome)
mr_df_chart$Outcome <- sub("^(.*)[.].*", "\\1",mr_df_chart$Outcome)

mr_df_chart <- mr_df_chart[!duplicated(mr_df_chart$Exposure), ]

library(scales)

ggplot(mr_df_chart, aes(x=reorder(as.factor(Exposure), Beta_exponent) , y=Beta_exponent)) + 
    geom_boxplot(fill="slateblue") + coord_flip() + geom_errorbar(aes(ymax = UpperCI, ymin = LowerCI))+ ggtitle("Outcome - Alzheimer's") +
  xlab("Exposure Variable") + ylab("Odds Ratio (95% Confidence Interval)")+ geom_hline(yintercept=1, linetype="dashed", color = "red") + 
  theme_bw() + theme(text=element_text(size=25, family="Arial")) +
  geom_hline(linetype = "dashed", yintercept = 1) +
    theme(panel.grid.major.x = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray41"),
        panel.grid.major.y = element_line(size = 0.1, 
                                        linetype = 'solid', 
                                        colour = "gray62")) +
  scale_y_log10(breaks = c(-1, 0.5, 1, 1.5, 2, 2.5)) +
  guides(color = guide_legend(reverse = TRUE)) +
  coord_flip(clip = "off", ylim = c(-1,3))

```

# AD -> PROTEIN

#SUN

```{r echo=FALSE}

#SUN

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR/sun_mr/ADexp/output/")

sun_list <- list.files(pattern = "*.gsmr")

sun_mr = lapply(sun_list, read.delim)

sun_mr <- do.call(rbind.data.frame, sun_mr)

sun_mr$study <- "Sun"

sun_mr$file_name <- rep(sun_list, each=2)

sun_mr <- sun_mr %>%
  filter(nsnp >= 1)

# sun_mr$Exposure <- sub("^(.*)[.].*", "\\1",sun_mr$Exposure)
# sun_mr$Exposure <- sub("^(.*)[.].*", "\\1",sun_mr$Exposure)
# sun_mr$Exposure <- sub("^(.*)[.].*", "\\1",sun_mr$Exposure)
# sun_mr$Exposure <- sub("^(.*)[.].*", "\\1",sun_mr$Exposure)
# 
# sun_mr$Outcome <- sub("^(.*)[.].*", "\\1",sun_mr$Outcome)
# sun_mr$Outcome <- sub("^(.*)[.].*", "\\1",sun_mr$Outcome)
# sun_mr$Outcome <- sub("^(.*)[.].*", "\\1",sun_mr$Outcome)
# sun_mr$Outcome <- sub("^(.*)[.].*", "\\1",sun_mr$Outcome)


sun_n <- count(sun_mr)
```

#SUHRE

```{r echo=FALSE}

#SUHRE
setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//suhr_mr/ADexp/outputs/")

suhre_list <- list.files(pattern = "*.gsmr")

suhre_mr = lapply(suhre_list, read.delim)

suhre_mr <- do.call(rbind.data.frame, suhre_mr)

suhre_mr$study <- "Suhre"

suhre_mr$file_name <- rep(suhre_list, each=2)

suhre_mr <- suhre_mr %>%
  filter(nsnp >= 1)
# 
# suhre_mr$Exposure <- gsub("\\..*","",suhre_mr$Exposure)
# suhre_mr$Exposure <- gsub("(.*)_.*", "\\1",suhre_mr$Exposure)
# suhre_mr$Exposure <- gsub("(.*)_.*", "\\1",suhre_mr$Exposure)
# suhre_mr$Exposure <- gsub("(.*)_.*", "\\1",suhre_mr$Exposure)
# suhre_mr$Exposure <- gsub("Suhre_Renamed/", "\\1",suhre_mr$Exposure)


suhre_n <- count(suhre_mr)
```

#FOLK 

```{r echo=FALSE}

#FOLK 

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//folk_mr/ADexp/output/")

folk_list <- list.files(pattern = "*.gsmr")

folk_mr = lapply(folk_list, read.delim)

folk_mr <- do.call(rbind.data.frame, folk_mr)

folk_mr$study <- "Folk"

folk_mr$file_name <- rep(folk_list, each=2)


folk_mr <- folk_mr %>%
  filter(nsnp >= 1)

folk_n <- count(folk_mr)

```

# #SLIZ 

```{r echo=FALSE}
# 
# #SLIZ 
# 
# setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//sliz_mr/ADexp/output/")
# 
# sliz_mr <- list.files(pattern = "*.gsmr")
# 
# sliz_mr = lapply(sliz_mr, read.delim)
# 
# sliz_mr <- do.call(rbind.data.frame, sliz_mr)
# 
# sliz_mr$study <- "Sliz"
# 
# 
# sliz_mr <- sliz_mr %>%
#   filter(nsnp >= 1)
# 
# sliz_n <- count(sliz_mr)
# 
```

#WOOD

```{r echo=FALSE}

#WOOD

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//wood_mr/ADexp/output/")

wood_list <- list.files(pattern = "*.gsmr")

wood_mr = lapply(wood_list, read.delim)

wood_mr <- do.call(rbind.data.frame, wood_mr)

wood_mr$study <- "Wood"

wood_mr$file_name <- rep(wood_list, each=2)


wood_mr <- wood_mr %>%
  filter(nsnp >= 1)

wood_n <- count(wood_mr)

```

#AHOL 

```{r echo=FALSE}

#AHOL 

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//infl_mr/ADexp/output/")

ahol_list <- list.files(pattern = "*.gsmr")

ahol_mr = lapply(ahol_list, read.delim)

ahol_mr <- do.call(rbind.data.frame, ahol_mr)

ahol_mr$study <- "Ahol"


ahol_mr$file_name <- rep(ahol_list, each=2)


ahol_mr <- ahol_mr %>%
  filter(nsnp >= 1)

ahol_n <- count(ahol_mr)

```

#SCALLOP

```{r echo=FALSE}

#SCALLOP

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//scal_mr/ADexp/output/")

scal_list <- list.files(pattern = "*.gsmr")

scal_mr = lapply(scal_list, read.delim)

scal_mr <- do.call(rbind.data.frame, scal_mr)

scal_mr$study <- "Scal"

scal_mr$file_name <- rep(scal_list, each=2)

scal_mr <- scal_mr %>%
  filter(nsnp >= 1)

scal_n <- count(scal_mr)

```

#HILL

```{r echo=FALSE}

#HILL

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//hill_mr/ADexp/output/")

hill_list <- list.files(pattern = "*.gsmr")

hill_mr = lapply(hill_list, read.delim)

hill_mr <- do.call(rbind.data.frame, hill_mr)

hill_mr$study <- "Hill"

hill_mr$file_name <- rep(hill_list, each=2)

hill_mr <- hill_mr %>%
  filter(nsnp >= 1)

hill_n <- count(hill_mr)

```

#HOGL

```{r echo=FALSE}

#HOGL

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//hogl_mr/ADexp/output/")

hogl_list <- list.files(pattern = "*.gsmr")

hogl_mr = lapply(hogl_list, read.delim)

hogl_mr <- do.call(rbind.data.frame, hogl_mr)

hogl_mr$study <- "Hogl"

hogl_mr$file_name <- rep(hogl_list, each=2)

hogl_mr <- hogl_mr %>%
  filter(nsnp >= 1)

hogl_n <- count(hogl_mr)

```

#ENROTH

```{r echo=FALSE}

#ENROTH

setwd("/scratch/groups/ukbiobank/usr/alish/AD_MR//enroth_mr//ADexp/output/")

enroth_list <- list.files(pattern = "*.gsmr")

enroth_mr = lapply(enroth_list, read.delim)

enroth_mr <- do.call(rbind.data.frame, enroth_mr)

enroth_mr$study <- "Enroth"

enroth_mr$file_name <- rep(enroth_list, each=2)

enroth_mr <- enroth_mr %>%
  filter(nsnp >= 1)

enroth_n <- count(enroth_mr)

```

#Count the number of anlayses done
```{r echo=FALSE}
total_tests <- sum(ahol_n$n, folk_n$n, scal_n$n, sliz_n$n, suhre_n$n, sun_n$n, wood_n$n, hill_n$n, hogl_n$n, enroth_n$n)

threshold <- 0.05/total_tests

threshold <- format(threshold,scientific=FALSE)

print(threshold)
```


# Combine all to count how many exposures we had as proteins
```{r echo=FALSE}
# combine all to count how many exposures and outcomes we had
mr_df_all <- rbind(folk_mr, suhre_mr, sun_mr, wood_mr, ahol_mr, scal_mr, hill_mr, hogl_mr, enroth_mr)
als <- str_count(mr_df_all$Exposure, "AD")
als[is.na(als)] <- 0
sum(als)
```

# Combine all to count how many exposures we had as ALS
```{r}
markers <- mr_df_all[- grep("AD_Schwartzentruber", mr_df_all$Outcome),]
nrow(markers)
```

# Combine full data, work out odds ratios and 95% CI
```{r}

mr_df_all$Beta_exponent <- exp(mr_df_all$bxy)
mr_df_all$LCI <- mr_df_all$Beta_exponent - (mr_df_all$se * 1.96)
mr_df_all$UCI <- mr_df_all$Beta_exponent + (mr_df_all$se * 1.96)

mr_df_all$Exposure <- gsub(".txt", "",mr_df_all$Exposure)
mr_df_all$Outcome <- gsub(".txt", "",mr_df_all$Outcome)
```

# Compute q-values and add them to the large combined data frame
```{r}
pvalues <- mr_df_all$p
qobj <- qvalue(p = pvalues)
lfdr <- qobj$lfdr
summary(qobj)
hist(qobj)
plot(qobj)

qvalues <- as.matrix(qobj$qvalues)
mr_df_all$qvalue <- cbind(qvalues)
```


# Adjusted p-values using SGoF method
```{r}
p <- SGoF(mr_df_all$p)
summary(p)

plot(p)

p_list <- p$Adjusted.pvalues

mr_df_all <- mr_df_all[order(mr_df_all$p),]
mr_df_all$sgof <- p_list
```

# All effects AD exposure
```{r}

mr_df_all$p.adjust <- p.adjust(p = mr_df_all$p, method = "fdr")

AD_exposure <- mr_df_all[!grepl("AD_Schwartzentruber", mr_df_all$Outcome),]

  AD_exposure %>%
  kbl() %>%
  kable_classic_2(full_width = F)

```

# Sort by significant p-values
```{r}
mr_df_filtered <- AD_exposure %>%
  filter(qvalue < 0.05) 

mr_df_filtered <- mr_df_filtered %>%
  rename(Beta = bxy, StandardError = se, Pvalue = p, SNPs = nsnp, LowerCI = LCI, UpperCI =  UCI, AdjustedPvalue = p.adjust)

mr_df_filtered <- mr_df_filtered %>% arrange(desc(Beta_exponent))
     
mr_df_filtered$Log_Beta <- log(mr_df_filtered$Beta_exponent) 
mr_df_filtered$Log_LowerCI <- log(mr_df_filtered$LowerCI) 
mr_df_filtered$Log_UpperCI <- log(mr_df_filtered$UpperCI) 
                                        
kable(mr_df_filtered)                                

```

# Create a figure with AD as the exposure 

```{r echo=FALSE}

mr_df_chart <- mr_df_filtered %>%
  filter(qvalue < 0.0001) 

mr_df_chart <- mr_df_chart[complete.cases(mr_df_chart), ]

#Tidy up the names of some proteins 
mr_df_chart$Exposure <- sub("^(.*)[.].*", "\\1",mr_df_chart$Exposure)
mr_df_chart$Exposure <- sub("^(.*)[.].*", "\\1",mr_df_chart$Exposure)
mr_df_chart$Exposure <- sub("^(.*)[.].*", "\\1",mr_df_chart$Exposure)
mr_df_chart$Exposure <- sub("^(.*)[.].*", "\\1",mr_df_chart$Exposure)

mr_df_chart$Outcome <- sub("^(.*)[.].*", "\\1",mr_df_chart$Outcome)
mr_df_chart$Outcome <- sub("^(.*)[.].*", "\\1",mr_df_chart$Outcome)
mr_df_chart$Outcome <- sub("^(.*)[.].*", "\\1",mr_df_chart$Outcome)
mr_df_chart$Outcome <- sub("^(.*)[.].*", "\\1",mr_df_chart$Outcome)

mr_df_chart <- mr_df_chart[!duplicated(mr_df_chart$Outcome), ]

library(scales)

ggplot(mr_df_chart, aes(x=reorder(as.factor(Outcome), Beta_exponent) , y=Beta_exponent)) + 
    geom_boxplot(fill="slateblue") + coord_flip() + geom_errorbar(aes(ymax = UpperCI, ymin = LowerCI))+ ggtitle("Exposure - Alzheimer's") +
  xlab("Outcome Variable") + ylab("Odds Ratio (95% Confidence Interval)")+ geom_hline(yintercept=1, linetype="dashed", color = "red") + ylim(-0.5, 2.5) + 
  theme_bw() + theme(text=element_text(size=10, family="Arial")) +
    theme(panel.grid.major.x = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray41"),
        panel.grid.major.y = element_line(size = 0.1, 
                                        linetype = 'solid', 
                                        colour = "gray62")) +
  scale_y_log10(breaks = c(-1, 0.5, 1, 1.5, 2, 2.5)) +
  guides(color = guide_legend(reverse = TRUE)) +
  coord_flip(clip = "off", ylim = c(-1,3))
  
```

# Get lists of proteins under 0.05 threshold for analysis with brain volumes 
```{r}

AD_exposure_list <- AD_exposure %>%
  filter(qvalue < 0.05) 


AD_outcome_list <- AD_outcome %>%
  filter(qvalue < 0.05) 

```


# Import Google sheet with all our marker info 
This is to get markers for the next stage of the anlayses
```{r}
  
#   library(gsheet)
# 
# #SUN
# 
# gsheet <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1SYa6zFg9ILj_R8YjBr01nYoEZ7-XGPKxpAWIUKlX3sE/edit#gid=776057591")
# 
# 
# idx2 <- sapply(AD_outcome_list$Exposure, grep, gsheet$file.name)
# 
# idx1 <- sapply(seq_along(idx2), function(i) rep(i, length(idx2[[i]])))
# 
# sun_match <- cbind(AD_outcome_list[unlist(idx1),,drop=F], gsheet[unlist(idx2),,drop=F])
# 
#   sun_match %>%
#   kbl() %>%
#   kable_classic_2(full_width = F)
```













